### 2023-02-28 paper/notes/notes.tex
notes

\documentclass[fleqn,11pt]{article}
\usepackage{amsmath, amssymb}
\title{notes}

\begin{document}
\subsection*{Notation}
Given a finite population $U=\{1,2,\ldots, N\}$ partitioned in domains
$\{D_i\}_{i=1}^M$, of sizes $N_i$.  Draw a sample $S\subseteq U$ of size
$n$ uniformly without replacement from $U$. We have

$$
\mathbb{P}(j\in S) = \frac{n}{N}\textrm{ for all } j\in U.
$$

Assume we measure a variable $Y$ with realisations $\{y_j|j\in S\}$. We shall
be interested in estimating the population total $t_Y$, the population mean
$\mu_Y$ and the totals and means for each domain, denoted $t_{Y|i}$ and $\mu_{Y|i}$
respectively.

\subsection*{Horwitz-Thomson estimation}
The Horwitz-Thomson estimator for the population total is defined as
$$
\hat{t}^{\textrm{HT}}_Y = \sum_{j\in S} \frac{1}{\mathbb{P}(j\in S)}y_j
=\sum_{j\in S}\frac{N}{n}y_j = N\bar{y},
$$
where $\bar{y}$ is the sample mean of $Y$.  The HT estimator for a domain $D_i$
is given by
$$
\hat{t}_{Y|i}^{\textrm{HT}} = \sum_{j\in S_i}\frac{N}{n}y_j,\textrm{ with } S_i=D_i\cap S.
$$
This estimator has the property that
$\sum_{i=1}^M\hat{t}_{Y|i}^{\textrm{HT}}=\hat{t}_{Y}^{\textrm{HT}}$, or more
generally, for any $L\subseteq \{1,2,\ldots,M\}$ we have the additivity
property $\hat{t}^{\textrm{HT}}_{Y|L} = \sum_{l\in
L}\hat{t}^{\textrm{HT}}_{Y|l}$.

For the corresponding estimator of the population mean we get the following.
$$
\hat{\mu}_{Y}^{\textrm{HT}} = \frac{1}{N}\hat{t}^{\textrm{HT}}_Y = 
\frac{1}{n}\sum_{j\in S}y_j = \bar{y}.
$$
For the domain means we define
$$
\hat{\mu}_{Y|i}^{\textrm{HT}} = \frac{1}{N_i}\hat{t}^{\textrm{HT}}_{Y|i}
= \frac{N}{N_i n}\sum_{j\in S_i}y_j.
$$
We can rewrite this by observing that $\mathbb{E}(n_i)=\frac{N_in}{N}$ to get
$$
\hat{\mu}_{Y|i}^{\textrm{HT}} = \frac{1}{\mathbb{E}(n_i)}\sum_{j\in S_i}y_j
=\frac{n_i}{\mathbb{E}(n_i)}\bar{y}_i,
$$
where $\bar{y}_i$ is the sample mean of $Y$ observed in $S_i$.

\subsection*{The BARE estimator for undersampled domains}
In cases where $n_i$ is small, the estimation variance may be unacceptably
large.  A variance-bias tradeoff can be made by including units outside $S_i$
in the estimates for $S_i$. This is called \emph{Small Area Estimation} (SAE).
In SAE jargon, this is called `borrowing statistical strength' from other
domains.

One of the simplest estimators is the Broader Area Ratio Estimator (BARE). The
idea is to estimate the total for a broader domain that includes $D_i$ and then
disaggregate the total proportional to the domain size $N_i$. Choosing the
whole population $U$ for the broader we get the following.
$$
\hat{t}^{\textrm{BARE}}_Y = \frac{N_i}{N}\hat{t}^{\textrm{HT}}_Y = N_i\bar{y}.
$$
(Note that $\hat{y}$ is the mean, taken over the whole sample $S$). And for the
mean:
$$
\hat{\mu}^{\textrm{BARE}}_Y  = \frac{1}{N_i}\hat{t}^{\textrm{BARE}}_Y  = \bar{y}.
$$
Thus, the BARE estimator for $\mu_{Y|i}$ for a domain $i$ is given by the
estimated mean of the population.

More generally, suppose that $L\subseteq \{1,2,\ldots,M\}$; choose a $D_i$
with $i\in L$, and write $S_L = \cup_{l\in L}{S_l}$, then
%
$$
\hat{t}^{\textrm{BARE}}_{Y|i}(L) 
= \frac{N_i}{N_L}\hat{t}^{\textrm{HT}}_{Y|L}
=\frac{N_i}{N_L}\sum_{j\in S_L}\frac{N}{n}y_l 
= \frac{N}{N_L n}N_i n_L\bar{y}_L.
$$ 
Noting furthermore that $\mathbb{E}(n_L)=N_Ln/N$ we get
$$
\hat{t}^{\textrm{BARE}}_{Y|i}(L)= N_i\frac{n_L}{\mathbb{E}(n_L)}\bar{y}_L,
= N_i \hat{\mu}^{\textrm{HT}}_{Y|L}.
$$
Finally, we get for the mean
$$
\hat{\mu}^{\textrm{BARE}}_{Y|i}(L) 
= \frac{1}{N_i}\hat{t}^{\textrm{BARE}}_{Y|i}(L)
= \hat{\mu}^{\textrm{HT}}_{Y|L}
$$

If we assume the approximation $n_L/\frac{\mathbb{E}(n_L}\approx 1$ we 
get
$$
\hat{\mu}^{\textrm{BARE}}_{Y|i}(L)\approx \bar{y}_L.
$$

\end{document}